Paper

### 2020-04-14 Paper.Rmd
Bug fixes for paper figures

all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
patient.info = all.patient.info

@@

```{r}

cohortSummary = sum.patient.data %>% dplyr::group_by(Status) %>% dplyr::summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  'Range endoscopies'=paste(range(total.endos), collapse='-'),
  'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Mean age'=paste(round(mean(age.diagnosis), 1),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':'),
  'Mean Length'=paste(round(mean(C,na.rm=T), 1),'+/-',round(sd(C,na.rm=T),1)),
  'naC'=length(which(is.na(C)))
)

cohortSummary = full_join(cohortSummary, cbind.data.frame('Status' = c('NP','P'), rbind(
      table(subset(patient.info, Status == 'NP')$Pathology),
      table(subset(patient.info, Status == 'P')$Pathology))), by='Status')

cohortSummary %>% t %>% kable(caption='Training set') %>% kable_styling()

# fit = glm(Status ~ age.diagnosis+M+gender, data=sum.patient.data, family = binomial(link='logit'))
# summary(fit)
# 
# sum.patient.data = sum.patient.data %>% mutate(total.followup = end.year - start.year)
# 
# surv =  Surv(sum.patient.data$total.followup,sum.patient.data$Status)
# terms = as.formula(paste("Surv(total.followup, status) ~", paste(c('age.diagnosis','M','gender'),collapse=" + "), sep=""))
# 
# cox = coxph(terms, data=sum.patient.data, na.action=na.omit,ties='breslow')
# summary(cox)
# 
# ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F, CI=T ) +
#       annotate("text", x=5, y=0.25,label=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),3),"\n", nrow(data), 'patients')) + theme_minimal()
# 
# 
# t.test(filter(sum.patient.data, Status == 'NP')$age.diagnosis,
#        filter(sum.patient.data, Status == 'P')$age.diagnosis)
# t.test(filter(sum.patient.data, Status == 'NP')$M, filter(sum.patient.data, Status == 'P')$M)
# fisher.test(table(sum.patient.data$Status, sum.patient.data$smoking))
# fisher.test(table(sum.patient.data$Status, sum.patient.data$gender))

@@

```{r, echo=F}
grid.arrange(
  ggplot(patient.info, aes(p53.Status, fill=Status)) + geom_bar(position = position_dodge()) + labs(title='p53 Staining') + plot.theme,
  tableGrob( 
    rbind('Non-progressor'=summary(subset(patient.info, Status == 'NP')$p53.Status), 
           'Progressor'=summary(subset(patient.info, Status == 'P')$p53.Status))), 
  nrow=2, as.table=T, heights=c(4,1) )
```

@@

#ddf = rbind(dysplasia.df, 'mean.value'=apply(dysplasia.df, 2, mean))
globalCNdf = as_tibble( t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))]), rownames = 'locs') %>% 
  separate(locs, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), as.numeric) %>%
  dplyr::mutate( seg.type = ifelse( (start %in% arm.loc$start & end %in% arm.loc$end), 'arms','5MB'  ) ) %>%
  melt(id.vars=c('locs','chr','start','end', 'seg.type')) %>% as_tibble() %>%
  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
  left_join(patient.info %>% dplyr::select(matches('Hospital|Patient|Samplename|Pathology|Status')), by=c('variable'='Samplename')) %>%
  dplyr::mutate(Status = factor(recode(Status, NP='Non-Progressor',P='Progressor'), ordered=T)) %>% 
  dplyr::mutate( 
    cn = factor(unlist(purrr::map(.x=value, .f=get.cn.state)),  levels=c('loss','neutral','gain'), ordered = T),
    chr = factor(chr, levels = chr.lengths$chrom, ordered=T))

@@

ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

@@

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))

ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)

```{r echo=F, fig.height=5, fig.width=5}

autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme

ggsave('plots/glm_0.9.png', plot=autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme, width=5, height=5, units='in', dpi=300)
```

@@

```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=T}
rm(performance.at.1se, coefs, mp)

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

nohgd.coefs = coefs[[select.alpha]]
#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
```

@@

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp)
```

@@

```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

#do.call(grid.arrange, c(nolgd.plots, top="No LGD/HGD/IMC samples", ncol=2))
```

@@

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp)
```

@@

```{r, echo=F, warning=F, message=F, fig.width=6, fig.height=6}
#classification.rate$x = 1:nrow(classification.rate)

p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
#  geom_point(size=2) +
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p

ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

@@

cvRR = features

qq = quantile(features$hazards, seq(0,1,.15))
p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by cv(RR)', x='Coefficient value', y='cvRR') + plot.theme 
p
ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

cvRR %>% mutate(`Gain/Loss` = ifelse(coef > 0 ,'Gain', 'Loss')) %>% 
  dplyr::select(label, hazards, `Gain/Loss`) %>% arrange(desc(hazards)) %>%
  write_tsv('plots/coef_effect.tsv')

`r features %>% kable(caption='Selected coefficients') %>% kable_styling(full_width=F)`

@@

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
# loo_residuals = do.call(bind_rows, purrr::map(list.files(paste0(dir, '/Analysis/pcf_perPatient/50kb'), 'residuals.tsv', recursive = T, full.names = T), function(f) {
#   read_tsv(f, col_types = 'ccddddl')
# }))
# 
# failedQC = loo_residuals %>% filter(varMAD_median > 0.008)

error = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/sample_error.tsv', col_types = 'cdD')

@@

#error %>% group_by(Status) %>% dplyr::summarise_at(vars(Abs.Error), list(~mean(.,na.rm=T), ~sd(.,na.rm=T)))
#summary(error$Abs.Error)
#sd(error$Abs.Error)

@@

pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE'))

pg.samp = pg.samp %>% left_join(error %>% dplyr::select(-RR,-Status), by=c('Samplename'='Sample'))

@@

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p
ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

#oddPt = names(pg.samp)[which.max(performance.at.1se)]
#subset(sum.patient.data, Hospital.Research.ID == oddPt)

#pg.samp[[oddPt]]$Prediction

# Not the pt with the most samples or endos
#sum.patient.data[which.max(sum.patient.data$total.samples),]
```

@@

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOHGD) + ggtitle('Model excluding HGD/IMC')

rocNOHGD$auc
pROC::ci.auc(rocNOHGD$auc)

ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD) + ggtitle('Model excluding HGD/IMC'), width=6, height=6, units='in', dpi=300)
```

@@

pROC::ci.auc(roc$auc)

ggsave(filename='plots/auc/training_auc.png', plot=roc.plot(roc), width=6.5, height=4.5, units='in', dpi=300)
ggsave(filename='plots/auc/training_auc_supp.png', plot=roc.plot(roc) + ggtitle('Per sample'), width=6, height=6, units='in', dpi=300)

@@

droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
droc
#ggsave(filename='plots/auc/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)

@@

#summary(filter(df, model == 'in-silico')$AUC)

blues = RColorBrewer::brewer.pal(5,'Blues')
p = ggplot(df, aes(Specificity, Sensitivity)) + 
    geom_line(color=blues[2]) + 
    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    scale_x_reverse() +
    geom_label(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('AUC\n',round(AUC,2))), color='black') +
    #scale_color_brewer(palette='Blues') +
    labs(title='ROC AUC', subtitle='in-silico splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none', text = element_text(size=14))
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

# Not sure why I did this
# df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
# thresholds = 
#   rbind.data.frame(
#     est=t(round(pROC::coords(roc, "best", transpose = T),2)),
#     max=t(round(pROC::coords(roc, 0.8, transpose = T),2)),
#     'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6, transpose = T),2)),
#     'All samples'=t(round(pROC::coords(roc, 0.6, transpose = T),2))
# )
# thresholds$label=rownames(thresholds)
# 
# ggplot() + 
#   geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() +
#   geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
#   geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
#     paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(4,'Dark2')) + scale_color_manual(values=brewer.pal(4,'Dark2')) +
#   labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

cutoff = round(pROC::coords(roc, 'best', ret = 'threshold'), 2)
```

@@

pROC::ci.auc(p53roc$auc)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')
ggsave('plots/auc/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC'), width=6, height=6, units="in", limitsize=F, dpi = 300)
```

@@

Based on HGD/IMC/LGD
```{r pathroc, echo=F, fig.height=4, fig.width=12 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

pathsamples = patient.info[,c('Pathology','Status')]
pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pROC::ci.auc(pathroc$auc)

pathroc$model = 'HGD & LGD'
ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'
ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)

grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)'), ncol=2)
```

@@

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
rocEndoM$model = 'Per Endoscopy Max'
ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

>>>>

all.patient.info = read.patient.info(patient.file, demo.file, set='All')
patient.info = bind_rows(all.patient.info$info, all.patient.info$replicates) # for the moment...

@@

```{r demo}

cohortSummary = sum.patient.data %>% dplyr::group_by(Status) %>% 
  dplyr::summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (dates)'=paste(min(start.year), max(end.year), sep='-'),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  #'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  #'Total Range endoscopies'=paste(range(total.endos), collapse='-'),
  #'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Median total endoscopies (IQR)'=paste0(median(total.endos), ' (', paste(quantile(total.endos, prob=c(0.25,0.75)), collapse='-'), ')'),
  'Median age at diagnosis (IQR)'=paste0(median(age.diagnosis), ' (', paste(quantile(age.diagnosis, prob=c(0.25,0.75)), collapse='-'), ')'),
  'Mean Length (not recorded)'=paste0(paste(round(mean(M,na.rm=T), 1),'+/-',round(sd(M,na.rm=T),1)), ' (', length( which(is.na(M) )),')'),
  'Gender F:M'=paste(table(gender),collapse=':'),
  'Smoking N:Y (not recorded)'=paste0(paste(table(smoking), collapse=':'), ' (', length(which(is.na(smoking) )), ')')
) %>% left_join(patient.info %>% group_by(Status, Pathology) %>% tally %>% spread(Pathology,n), by='Status')

cohortSummary = cohortSummary %>% add_row(
  Status = 'p-value (test)', 'Total Patients'= '', 'Total Samples' = '', 'Range follow-up (dates)' = '', 
  'Range follow-up (years)' = paste0(signif(do(sum.patient.data, tidy(wilcox.test(end.year-start.year ~ Status, data=., exact=F)))$p.value,3), ' (Wilcox)'),
  'Median age at diagnosis (IQR)' = paste0(signif(do(sum.patient.data, tidy(wilcox.test(age.diagnosis ~ Status, data = ., exact=F)))$p.value,3), ' (Wilcox)'),
  'Mean Length (not recorded)' = paste0(signif(do(sum.patient.data, tidy(wilcox.test(M ~ Status, data=., exact=F)))$p.value,3), ' (Wilcox)'),
  'Gender F:M' = paste0(signif(dplyr::summarise(sum.patient.data, pvalue = chisq.test(gender, Status)$p.value)$pvalue,3), ' (ChiSq)'),
  'Smoking N:Y (not recorded)' = paste0(signif(dplyr::summarise(sum.patient.data, pvalue = chisq.test(smoking, Status)$p.value)$pvalue,3), ' (ChiSq)')
)

cohortSummary %>% t %>% kable(caption='Training Cohort') %>% kable_styling()

fit = glm(Status ~ age.diagnosis+M+gender+initial.path, data=mutate(sum.patient.data, initial.path = factor(initial.path,ordered=F)), family = binomial(link='logit'))
tidy(fit) %>% kable(caption = 'Logistic model to test effects of (matched) demographic factors on progression') %>% 
  kable_styling(full_width = F)

sample.age = patient.info %>% group_by(Samplename, Status, Pathology) %>% dplyr::summarise(sample.age = 2020-Endoscopy.Year)

ggplot(sample.age, aes(Status, sample.age, group=Status, fill=Status)) + geom_jitter(width=0.2) + geom_boxplot(alpha =0.5, width=0.5, outlier.color = NA) + 
  scale_y_continuous(breaks = seq(0,28,5)) + labs(y='Years', title='Sample age') + plot.theme

ggplot(sample.age, aes(Pathology, sample.age, group=Pathology, fill=Pathology, color=Status)) + 
  geom_jitter(width=0.2) + geom_boxplot(width=0.5, outlier.color = NA, color='black') + 
  scale_y_continuous(breaks = seq(0,28,5)) + labs(y='Years', title='Sample age') + plot.theme

@@

```{r, echo=F, eval=F}
grid.arrange(
  ggplot(patient.info, aes(p53.Status, fill=Status)) + geom_bar(position = position_dodge()) + labs(title='p53 Staining') + plot.theme,
  tableGrob( 
    patient.info %>% mutate(p53.Status = forcats::fct_explicit_na(p53.Status)) %>% group_by(Status, p53.Status) %>% tally() %>% spread(p53.Status, n)),
  nrow=2, as.table=T, heights=c(4,1) )  
```

@@

globalCNdf = as_tibble( t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))]), rownames = 'locs') %>% 
  separate(locs, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), as.numeric) %>%
  dplyr::mutate( seg.type = ifelse( (start %in% arm.loc$start & end %in% arm.loc$end), 'arms','5MB'  ) ) %>%
  melt(id.vars=c('locs','chr','start','end', 'seg.type')) %>% as_tibble() %>%
  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
  left_join(patient.info %>% dplyr::select(matches('Hospital|Patient|Samplename|Pathology|Status')), by=c('variable'='Samplename')) %>%
  dplyr::mutate(Status = factor(recode(Status, NP='Non-Progressor',P='Progressor'), ordered=T)) %>% 
  dplyr::mutate( 
    cn = factor(unlist(purrr::map(.x=value, .f=get.cn.state)),  levels=c('loss','neutral','gain'), ordered = T),
    chr = factor(chr, levels = chr.lengths$chrom, ordered=T))

globalCNdf %<>% filter(!is.na(Status))

@@

@@

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))
ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)
```

```{r echo=F, fig.height=5, fig.width=5}
autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme
ggsave('plots/glm_0.9.png', plot=autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme, width=5, height=5, units='in', dpi=300)
```

@@

```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=T}
rm(performance.at.1se, coefs, mp)
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

nohgd.coefs = coefs[[select.alpha]]
ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
```

@@

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
rm(performance.at.1se, coefs, mp)
```

@@

```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}
```

@@

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
rm(performance.at.1se, coefs, mp)
```

@@

```{r, echo=F, warning=F, message=F, fig.width=6, fig.height=6}
p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
#  geom_point(size=2) +
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p
ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

@@

cvRR = features
qq = quantile(features$hazards, seq(0,1,.15))
p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by cv(RR)', x='Coefficient value', y='cvRR') + plot.theme 
p
ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

cvRR %>% mutate(`Gain/Loss` = ifelse(coef > 0 ,'Gain', 'Loss')) %>% 
  dplyr::select(label, hazards, `Gain/Loss`) %>% arrange(desc(hazards)) %>%
  write_tsv('plots/coef_effect.tsv')
```

`r features %>% arrange(desc(hazards)) %>% kable(caption='Selected coefficients') %>% kable_styling(full_width=F)`

@@

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
error = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/sample_error.tsv', col_types = 'cdD')

@@

@@

pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE'))
pg.samp = pg.samp %>% left_join(error %>% dplyr::select(-RR,-Status), by=c('Samplename'='Sample'))

@@

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p
ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
```

@@

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T)

pROC::coords(rocNOHGD, 'best', transpose=F) %>% add_column('AUC'=pROC::auc(rocNOHGD)) %>% mutate_all(list(~round(.,2))) %>% 
  kable(caption = 'LOO excluding HGD/IMC') %>% kable_styling(full_width = F)

ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T) , width=6, height=6, units='in', dpi=300)
```

#### LOO minus HGD/LGD

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonolgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}

#length(performance.at.1se)
#length(unique(pg.sampNOLGD$Patient))

#perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOLGD$Patient))

pg.sampNOLGD = pg.sampNOLGD %>% filter(!is.na(Prediction))

rocNOLGD = pROC::roc(Status ~ Prediction, data=pg.sampNOLGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',incS = T)

pROC::coords(rocNOLGD, 'best', transpose=F) %>% add_column('AUC'=pROC::auc(rocNOLGD)) %>% mutate_all(list(~round(.,2))) %>% 
  kable(caption = 'LOO excluding LGD/HGD/IMC') %>% kable_styling(full_width = F)

ggsave(filename='plots/auc/roc_nolgd_model.png', plot=roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',incS = T) , width=6, height=6, units='in', dpi=300)
```

```{r}
roc$model = 'All samples model'
rocNOHGD$model = 'Excluding HGD/IMC'
rocNOLGD$model = 'Excluding LGD/HGD/IMC'

m = bind_rows(
      get.roc.stat(roc),
      get.roc.stat(rocNOHGD)
#      get.roc.stat(rocNOLGD)
      ) %>% dplyr::mutate(model = factor(model)) %>% dplyr::mutate_if(is.numeric, list(~round(.,2)))

p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
  geom_errorbar(position=dodge, width=0.3, color='black' ) +
  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
  scale_y_continuous(limits=c(0,1)) +
  scale_fill_brewer(palette='Set2') +
  theme(legend.position = 'right',text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
  labs(x='', y='ROC', title='Model comparisons')

auc.leg = get.legend(p)

ggsave('plots/auc/loo_model_roc.png', plot=p + theme(legend.position = 'none'), height=6, width=4.5, units="in", limitsize=F, dpi = 300)

#ggsave('plots/auc/multi_auc.png', plot=p+theme(legend.position = 'none'), height=6, width=8, units='in',dpi=300)
```

#### Insilico splits

@@

roc.plot(roc,'Discovery Cohort all samples ROC',incS=T)
round(pROC::coords(roc, 'best', transpose=F),2) %>% kable(caption = 'All samples ROC') %>% kable_styling(full_width = F)

#pROC::coords(roc, 0.3, input='threshold', transpose=T)

ggsave(filename='plots/auc/training_auc.png', plot=roc.plot(roc,'Discovery Cohort all samples ROC',incS = T),width=6, height=6, units='in', dpi=300)
ggsave(filename='plots/auc/training_auc_supp.png', plot=roc.plot(roc,'Discovery Cohort all samples ROC'), width=6, height=6, units='in', dpi=300)

@@

@@

blues = RColorBrewer::brewer.pal(5,'Greys')
p = ggplot(df, aes(Specificity, Sensitivity)) + 
    geom_line(color=blues[2]) + 
    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    scale_x_reverse() +
    geom_label_repel(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('FPR ', round(1-Specificity,2), ' TPR ', round(Sensitivity,2), '\nAUC ',round(AUC,2))), color='black') +
    #scale_color_brewer(palette='Blues') +
    labs(title='Discovery cohort: 50 models', subtitle='18:70 random patient splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none', text = element_text(size=14))
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
```

@@

#pROC::ci.auc(p53roc$auc)
#pROC::coords(p53roc, 'best',transpose=T)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC', incS = T) 
ggsave('plots/auc/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC', incS = T) , width=6, height=6, units="in", limitsize=F, dpi = 300)
```

@@

Based on HGD/IMC/LGD
```{r pathroc, echo=F, fig.height=4, fig.width=12 }
pathsamples = patient.info[,c('Pathology','Status')]
pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
#pROC::ci.auc(pathroc$auc)
#pROC::coords(pathroc, 'best')

pathroc$model = 'HGD & LGD'
ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)', incS = T), width=6, height=6, units="in", limitsize=F, dpi = 300)

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )

hgdpathroc$model = 'HGD Only'
ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD only', incS = T), width=6, height=6, units="in", limitsize=F, dpi = 300)

grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)', incS = T), ncol=2)
```

@@

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
rocEndoM$model = 'Max per Endoscopy'
ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
